Fully Automatic Loop-free Program Synthesis by Bounded Model Checking

### 2013-10-09 paper/paper.tex
Word width parameter

@@

\section{Technical Contributions}
Our approach generates an $\exists \forall$ formula which is linear
in the size of the synthesised program.  This improves over the
enoding of~\cite{brahma}, which generates a $\exists \forall$ formula
that is quadratic in the size of the synthesised program.

@@

It is often the case that a program which satisfies the specification will
continue to satisfy it regardless of the word width of the machine it is
implemented on.  For example, the program in Fig.~\ref{fig:bitvector-program}
isolates the least-significant bit of an 8-bit word, but it will also
isolate the least-significant bit of a 32-bit word or indeed of a word of
any size.

Because of this, we can often find a program that is correct for a large
word size by synthesizing a program for a machine with a smaller word size.
This reduces the size of all the constants in the program, the search space
of possible test inputs and also the size of all the state variables internal
to the $\mathcal{L}$-interpreter.

The only wrinkle here is that we will sometimes synthesise a program containing
constants.  If we have synthesised a program for a machine with $k$-bit words,
the constants in the program will be $k$ bits wide.  To generalize the program
to an $n$-bit machine (with $n > k$), we need some way of deriving $n$-bit-wide
numbers from $k$-bit ones.  We have several strategies for this and
just try each in turn.  In the following, a subscript $n$ on a value means that
the value is an $n$-bit word.  The dot operator $a_l \cdot b_m$ denotes
concatenation, resulting in an $(l+m)$-bit wide word.  A superscript on a value
means duplication, so $a_l^m$ means concatenating together $m$ copies of the $l$-bit
word $a$, resulting in an $(lm)$-bit wide word.

To generalize a number $x$ we use the following rules:

\begin{itemize}
 \item If $x = k_k$, generalize to $n_n$.
 \item If $x = k-1_k$, generalize to $n-1_n$.
 \item If $x = k+1_k$, generalize to $n+1_n$.
 \item Generalize $x_k$ to $x_n$.
 \item Generalize $x_k$ to $x_k \cdot 0_{n-k}$.
 \item Generalize $x_k$ to $x_k^{n / k}$.
\end{itemize}

>>>>

\newcommand{\bv}[2]{\mathcal{BV}(#1, #2)}

@@

\section{Technical Contributions}
Our approach generates an $\exists \forall$ formula which is linear
in the size of the synthesised program.  This improves over the
encoding of~\cite{brahma}, which generates a $\exists \forall$ formula
that is quadratic in the size of the synthesised program.

@@

It is often the case that a program which satisfies the specification
on an $\mathcal{L}$-machine with $w = k$ will continue to satisfy the
specification when run on a machine with $w > k$.  For example, the program
in Fig.~\ref{fig:bitvector-program} isolates the least-significant bit of a word.
This is true irrespective of the word size of the machine it is run on -- it will
isolate the least-significant bit of an 8-bit word just as well as it will a
32-bit word.  An often successful strategy is to synthesise a program for an
$\mathcal{L}$-machine with a small word size and then to check whether the
same program is correct when run on an $\mathcal{L}$-machine with a full
sized word.

The only wrinkle here is that we will sometimes synthesise a program containing
constants.  If we have synthesised a program for with $w=k$,
the constants in the program will be $k$ bits wide.  To generalize the program
to an $n$-bit machine (with $n > k$), we need some way of deriving $n$-bit-wide
numbers from $k$-bit ones.  We have several strategies for this and
just try each in turn.  The strategies are shown in Fig.~\ref{fig:generalize},
where $\mathcal{BV}(v, n)$ denotes an $n$-bit wide bitvector holding the value $v$
and $b \cdotp c$ means the concatenation of bitvectors $b$ and $c$.

\begin{figure}
To generalize an $m$-bit wide number to an $n$-bit wide one:

\begin{eqnarray*}
 \bv{m}{m} & \rightarrow & \bv{n}{n} \\
 \bv{m-1}{m} & \rightarrow & \bv{n-1}{n} \\
 \bv{m+1}{m} & \rightarrow & \bv{n+1}{n} \\
 \bv{x}{m} & \rightarrow & \bv{x}{n} \\
 \bv{x}{m} & \rightarrow & \bv{x}{m} \cdotp \bv{0}{n - m} \\
 \bv{x}{m} & \rightarrow & \underbrace{\bv{x}{m} \cdotp \ldots \cdotp \bv{x}{m}}_{\frac{n}{m} \mathrm{ times}}
\end{eqnarray*}

\caption{Rules for generalizing program constants
 \label{fig:generalize}}
\end{figure}
